---
title: "mCTB_in_R"
date: "April 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(multiwayvcov)
library(broom)
```
EXPAND DATASET TO 1 OBSERVATION PER BUDGETxINDIVIDUAL, MERGE IN INSTRUMENT DETAILS
```{r data setup}
data = read.table("/Users/rx/Desktop/mCTB Kit/data.txt", header=TRUE)
longdata=reshape(data, 
                 varying= c("c1","c2","c3","c4","c5","c6","c7","c8","c9","c10","c11","c12","c13","c14","c15","c16","c17","c18","c19","c20","c21","c22","c23","c24"), v.names="c",timevar="budget_number", times=c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23","24"), direction="long")
design=read.table("/Users/rx/Desktop/mCTB Kit/instrument_details.txt", header=TRUE)
finaldata = merge(longdata,design, by="budget_number", all.x=T, all.y=T)
```
CREATE CHOICESET VARIABLES
```{r choiset variable creation}
finaldata["t0"] = 0
finaldata$t0[finaldata$sooner_date == 0] = 1
finaldata$t0[finaldata$sooner_date != 0] = 0
finaldata$k=7*as.vector(finaldata$delay_weeks)
finaldata$pratio=(finaldata$endowment_later/finaldata$endowment_soon)
```
DEFINE VALUES FOR EACH OPTION
```{r value definition}
for (j in 1:6){
  finaldata[paste('soon_', j, sep='')]=(20/as.vector(finaldata$pratio))-(j-1)*(20/as.vector(finaldata$pratio))/5
}

finaldata$soon_6=0

for (j in 1:6){
  finaldata[paste('late_', j, sep='')]=(j-1)*4
}
```
ASSOCIATE CHOICE OPTIONS WITH VALUES
```{r choice association}
for (j in 1:6){
  for (i in 1:nrow(finaldata)){
    if (finaldata$c[i]==j){
      finaldata$temp=eval(as.name(paste('soon_', j, sep='')),finaldata)
      finaldata$sooner_choice[i] =finaldata$temp[i]
    }
  }
}
```
USE NON-LINEAR LEAST SQUARES TO ESTIMATE AGGREGATE PARAMETERS OF BETA-DELTA, CRRA, TIME-SEPARABLE UTILITY
```{r non-linear least squares aggregate parameter estimation}
a_start=0.90
b_start=1
d_start=0.999
model=nls(sooner_choice~endowment_later*((((b^t0)*(d^k)*pratio)^(1/(a-1)))/(1+pratio*((b^t0*d^k)*pratio)^(1/(a-1)))), finaldata, start=list(a=a_start,b=b_start, d=d_start))
summary(model, digits=4)
###IMPORTANT: THE STATA MODEL USES CLUSTERED SEs WHILE R DOES NOT###
vcov_subj = cluster.vcov(model, finaldata$subject_id)
coeftest(model, vcov_subj)
##STORE MODEL PARAMS##
finaldata$alpha=summary(model)$coefficients[1]
finaldata$beta=summary(model)$coefficients[2]
finaldata$delta=summary(model)$coefficients[3]
```
TRANSFORM DISCOUNT FACTOR TO MONTHLY DISCOUNT RATE
```{r monthly discount rate}
rate = delta^(-30) - 1
```
USE NON-LINEAR LEAST SQUARES TO ESTIMATE INDIVIDUAL PARAMETERS OF BETA-DELTA, CRRA, TIME-SEPARABLE UTILITY (MAY TAKE SOME TIME) GARBAGE BELOW 
```{r non-linear least squares individual parameter estimation }
finaldata$alpha_ind=NA
finaldata$beta_ind=NA
finaldata$delta_ind=NA
id=group_by(finaldata, subject_id)

for (i in 1:length(unique(id$subject_id))){
id %>% group_by(subject_id) %>%  try(do(tidy(nls(sooner_choice~endowment_later*((((b^t0)*(d^k)*pratio)^(1/(a-1)))/(1+pratio*((b^t0*d^k)*pratio)^(1/(a-1)))), finaldata, start=list(a=a_start,b=b_start, d=d_start), data=.)))) 
  finaldata$alpha_ind[i]=summary(tempmodel)$coefficients[1]
  finaldata$beta_ind[i]=summary(tempmodel)$coefficients[2]
  finaldata$delta_ind[i]=summary(tempmodel)$coefficients[3]
}

for (i in 1:length(unique(id$subject_id))){
  if (id$subject_id[i] == i){
    tempmodel=nls(sooner_choice~endowment_later*((((b^t0)*(d^k)*pratio)^(1/(a-1)))/(1+pratio*((b^t0*d^k)*pratio)^(1/(a-1)))), finaldata, start=list(a=a_start,b=b_start, d=d_start))
  finaldata$alpha_ind[i]=summary(tempmodel)$coefficients[1]
  finaldata$beta_ind[i]=summary(tempmodel)$coefficients[2]
  finaldata$delta_ind[i]=summary(tempmodel)$coefficients[3]
  }
}

for (i in 1:length(unique(id$subject_id))){
  if (id$subject_id[i] == i){
    testing <- id %>% group_by(subject_id) %>%
    do(tidy(nls(sooner_choice~endowment_later*((((b^t0)*(d^k)*pratio)^(1/(a-1)))/(1+pratio*((b^t0*d^k)*pratio)^(1/(a-1)))), finaldata, start=list(a=a_start,b=b_start, d=d_start))))
  finaldata$alpha_ind[finaldata$subject_id == i] = summary(tempmodel)$coefficients[1]
  finaldata$beta_ind[finaldata$subject_id == i] = summary(tempmodel)$coefficients[2]
  finaldata$delta_ind[finaldata$subject_id == i] = summary(tempmodel)$coefficients[3]
  }
}


 tempmodel=nls(sooner_choice~endowment_later*((((b^t0)*(d^k)*pratio)^(1/(a-1)))/(1+pratio*((b^t0*d^k)*pratio)^(1/(a-1)))), finaldata, start=list(a=a_start,b=b_start, d=d_start))


summarise(id, mean(sooner_choice))



