---
title: "mCTB_in_R"
description: "Conversion of the Time Preference Elicitation script written for STATA by Dr. Michael A. Kuhn into R. The mCTB is from Andreoni, Kuhn and Sprenger (2015)."
date: "April 18, 2017"
author: "Rita M Ludwig"
output: html_document
---

```{r setup, include=FALSE}
library(nlme)
library(dplyr)
library(multiwayvcov)
library(broom)
```
EXPAND DATASET TO 1 OBSERVATION PER BUDGETxINDIVIDUAL, MERGE IN INSTRUMENT DETAILS
```{r data setup}
data = read.table("/home/jflournoy/code_new/poverty/mCTB_kit/data.txt", header=TRUE)
longdata=reshape(data, 
                 varying= c("c1","c2","c3","c4","c5","c6","c7","c8","c9","c10","c11","c12","c13","c14","c15","c16","c17","c18","c19","c20","c21","c22","c23","c24"), v.names="c",timevar="budget_number", times=c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23","24"), direction="long")
design=read.table("/home/jflournoy/code_new/poverty/mCTB_kit/instrument_details.txt", header=TRUE)
finaldata = merge(longdata,design, by="budget_number", all.x=T, all.y=T)
```
CREATE CHOICESET VARIABLES
```{r choiset variable creation}
finaldata["t0"] = 0
finaldata$t0[finaldata$sooner_date == 0] = 1
finaldata$t0[finaldata$sooner_date != 0] = 0
finaldata$k=7*as.vector(finaldata$delay_weeks)
finaldata$pratio=(finaldata$endowment_later/finaldata$endowment_soon)
```
DEFINE VALUES FOR EACH OPTION
```{r value definition}
for (j in 1:6){
  finaldata[paste('soon_', j, sep='')]=(20/as.vector(finaldata$pratio))-(j-1)*(20/as.vector(finaldata$pratio))/5
}

finaldata$soon_6=0

for (j in 1:6){
  finaldata[paste('late_', j, sep='')]=(j-1)*4
}
```
ASSOCIATE CHOICE OPTIONS WITH VALUES
```{r choice association}
for (j in 1:6){
  for (i in 1:nrow(finaldata)){
    if (finaldata$c[i]==j){
      finaldata$temp=eval(as.name(paste('soon_', j, sep='')),finaldata)
      finaldata$sooner_choice[i] =finaldata$temp[i]
    }
  }
}
```
USE NON-LINEAR LEAST SQUARES TO ESTIMATE AGGREGATE PARAMETERS OF BETA-DELTA, CRRA, TIME-SEPARABLE UTILITY
```{r non-linear least squares aggregate parameter estimation}
a_start=0.90
b_start=1
d_start=0.999
model=nls(sooner_choice~endowment_later*((((b^t0)*(d^k)*pratio)^(1/(a-1)))/(1+pratio*((b^t0*d^k)*pratio)^(1/(a-1)))), finaldata, start=list(a=a_start,b=b_start, d=d_start))
summary(model, digits=4)
###IMPORTANT: THE STATA MODEL USES CLUSTERED SEs WHILE R DOES NOT###
#vcov_subj = cluster.vcov(model, finaldata$subject_id)
#coeftest(model, vcov_subj)
##STORE MODEL PARAMS##
finaldata$alpha=summary(model)$coefficients[1]
finaldata$beta=summary(model)$coefficients[2]
finaldata$delta=summary(model)$coefficients[3]
```
TRANSFORM DISCOUNT FACTOR TO MONTHLY DISCOUNT RATE
```{r monthly discount rate}
rate = finaldata$delta^(-30) - 1
```
USE NON-LINEAR LEAST SQUARES TO ESTIMATE INDIVIDUAL PARAMETERS OF BETA-DELTA, CRRA, TIME-SEPARABLE UTILITY (MAY TAKE SOME TIME) GARBAGE BELOW 
```{r non-linear least squares individual parameter estimation }
finaldata$alpha_ind=NA
finaldata$beta_ind=NA
finaldata$delta_ind=NA
id=group_by(finaldata, subject_id)
##EXAMPLE OF WHAT THE MODEL SHOULD LOOK LIKE##
examplemodel=nls(sooner_choice~endowment_later*((((b^t0)*(d^k)*pratio)^(1/(a-1)))/(1+pratio*((b^t0*d^k)*pratio)^(1/(a-1)))), filter(finaldata, subject_id==1), start=list(a=a_start,b=b_start, d=d_start))
##RUN THE INDIVIDUAL NLS MODELS##
individnlscoef=finaldata %>% group_by(subject_id)  %>% 
  do({
    tempmodel=try(nls(sooner_choice~endowment_later*((((b^t0)*(d^k)*pratio)^(1/(a-1)))/(1+pratio*((b^t0*d^k)*pratio)^(1/(a-1)))), ., start=list(a=a_start,b=b_start, d=d_start)))
  if(class(tempmodel)=='try-error'){
    returned=data.frame()
  }else{
    returned=data.frame(
      alpha_ind=summary(tempmodel)$coefficients[1],
      beta_ind=summary(tempmodel)$coefficients[2],
      delta_ind=summary(tempmodel)$coefficients[3]
    )
  }
  return(returned)
  })
##JOIN THE COEFFICIENT DATAFRAME WITH THE DATASET##
finaldata=left_join(finaldata, individnlscoef, by="subject_id")
```
TRANSFORM DISCOUNT FACTOR TO ANNUAL DISCOUNT RATE
```{r non-linear least squares individual parameter estimation }
finaldata$rate_ind = finaldata$delta_ind^(-30) - 1
```

# Try out nlme

An example:

```{r}
str(Orange)
GGally::ggpairs(Orange)

summary(Orange$circumference)

nlMLMod <- nlme(circumference ~ phi1 / (1 + exp(-(age - phi2)/phi3)),
                data = Orange,
                fixed = phi1 + phi2 + phi3 ~ 1,
                random = pdDiag(phi1 + phi2 ~ 1),
                groups = ~ Tree,
                start = c(phi1 = 200, phi2 = 700, phi3 = 350))

summary(nlMLMod)
coef(nlMLMod)
```


```{r}
a_start=0.90
b_start=1
d_start=0.999

#scale the outcome
finaldata$sooner_choice_scaled <- scale(finaldata$sooner_choice)

choice_nlme_mod=nlme(sooner_choice_scaled~endowment_later*((((b^t0)*(d^k)*pratio)^(1/(a-1)))/(1+pratio*((b^t0*d^k)*pratio)^(1/(a-1)))), 
              data = finaldata, 
              fixed = a + b + d ~ 1,
              random = pdDiag(a + b + d ~ 1),
              groups = ~ subject_id,
              start = c(a=a_start,b=b_start, d=d_start))

plot(ranef(choid_nlme_mod))

```