---
title: "mCTB_in_R"
description: "Conversion of the Time Preference Elicitation script written for STATA by Dr. Michael A. Kuhn into R. The mCTB is from Andreoni, Kuhn and Sprenger (2015)."
date: "April 18, 2017"
author: "Rita M Ludwig & John C Flournoy"
output:
  html_document:
    self_contained: no
---

To be clear about the model we're fitting:

$$
\text{sooner_choice}=\frac{\text{endowment_later}\cdot(b^{t_0}d^k\text{pratio})^{\frac{1}{a-1}}}{1+\text{pratio}\cdot(b^{t_0}d^k\text{pratio})^{\frac{1}{a-1}}}
$$
where $\text{endowment_later}=20$.

In R, this will look like:

```
sooner_choice ~ ( endowment_later * (b^t0 * d^k * pratio)^(1/(a-1)) ) / ( 1+pratio * (b^t0 * d^k * pratio)^(1/(a-1)) )
```

```{r setup, include=FALSE}
library(nlme)
library(dplyr)
library(multiwayvcov)
library(broom)
```
EXPAND DATASET TO 1 OBSERVATION PER BUDGETxINDIVIDUAL, MERGE IN INSTRUMENT DETAILS
```{r data setup}
data = read.table("/home/jflournoy/code_new/poverty/mCTB_kit/data.txt", header=TRUE)
longdata=reshape(data, 
                 varying= c("c1","c2","c3","c4","c5","c6","c7","c8","c9","c10","c11","c12","c13","c14","c15","c16","c17","c18","c19","c20","c21","c22","c23","c24"), v.names="c",timevar="budget_number", times=c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23","24"), direction="long")
design=read.table("/home/jflournoy/code_new/poverty/mCTB_kit/instrument_details.txt", header=TRUE)
finaldata = merge(longdata,design, by="budget_number", all.x=T, all.y=T)
```
CREATE CHOICESET VARIABLES
```{r choiset variable creation}
finaldata["t0"] = 0
finaldata$t0[finaldata$sooner_date == 0] = 1
finaldata$t0[finaldata$sooner_date != 0] = 0
finaldata$k=7*as.vector(finaldata$delay_weeks)
finaldata$pratio=(finaldata$endowment_later/finaldata$endowment_soon)
```
DEFINE VALUES FOR EACH OPTION
```{r value definition}
for (j in 1:6){
  finaldata[paste('soon_', j, sep='')]=(20/as.vector(finaldata$pratio))-(j-1)*(20/as.vector(finaldata$pratio))/5
}

finaldata$soon_6=0

for (j in 1:6){
  finaldata[paste('late_', j, sep='')]=(j-1)*4
}
```
ASSOCIATE CHOICE OPTIONS WITH VALUES
```{r choice association}
for (j in 1:6){
  for (i in 1:nrow(finaldata)){
    if (finaldata$c[i]==j){
      finaldata$temp=eval(as.name(paste('soon_', j, sep='')),finaldata)
      finaldata$sooner_choice[i] =finaldata$temp[i]
    }
  }
}
```
USE NON-LINEAR LEAST SQUARES TO ESTIMATE AGGREGATE PARAMETERS OF BETA-DELTA, CRRA, TIME-SEPARABLE UTILITY
```{r non-linear least squares aggregate parameter estimation one}
sd_of_soonerchoice <- sd(finaldata$sooner_choice)

#scale the outcome
finaldata$sooner_choice_scaled <- finaldata$sooner_choice / sd_of_soonerchoice
finaldata$endowment_later_scaled <- finaldata$endowment_later / sd_of_soonerchoice

a_start=0.90
b_start=1
d_start=0.999
model=nls(sooner_choice ~ ( endowment_later * (b^t0 * d^k * pratio)^(1/(a-1)) ) / ( 1+pratio * (b^t0 * d^k * pratio)^(1/(a-1)) ),
          finaldata, start=list(a=a_start,b=b_start, d=d_start))
summary(model, digits=4)

###IMPORTANT: THE STATA MODEL USES CLUSTERED SEs WHILE R DOES NOT###
#vcov_subj = cluster.vcov(model, finaldata$subject_id)
#coeftest(model, vcov_subj)
##STORE MODEL PARAMS##
finaldata$alpha=summary(model)$coefficients[1]
finaldata$beta=summary(model)$coefficients[2]
finaldata$delta=summary(model)$coefficients[3]
```
TRANSFORM DISCOUNT FACTOR TO MONTHLY DISCOUNT RATE
```{r monthly discount rate}
rate = unique(finaldata$delta^(-30) - 1)
```

USE NON-LINEAR LEAST SQUARES TO ESTIMATE INDIVIDUAL PARAMETERS OF BETA-DELTA, CRRA, TIME-SEPARABLE UTILITY (MAY TAKE SOME TIME) 

```{r non-linear least squares individual parameter estimation two, message=F, warning=F}
id=group_by(finaldata, subject_id)
##EXAMPLE OF WHAT THE MODEL SHOULD LOOK LIKE##
examplemodel=nls(sooner_choice ~ ( endowment_later * (b^t0 * d^k * pratio)^(1/(a-1)) ) / ( 1+pratio * (b^t0 * d^k * pratio)^(1/(a-1)) ), 
                 filter(finaldata, subject_id==1), start=list(a=a_start,b=b_start, d=d_start))
##RUN THE INDIVIDUAL NLS MODELS##
individnlscoef=finaldata %>% group_by(subject_id)  %>% 
  do({
    tempmodel=try(nls(sooner_choice ~ ( endowment_later * (b^t0 * d^k * pratio)^(1/(a-1)) ) / ( 1+pratio * (b^t0 * d^k * pratio)^(1/(a-1)) ),
                      ., 
                      start=list(a=a_start,b=b_start, d=d_start)))
  if(class(tempmodel)=='try-error'){
    returned=data.frame()
  }else{
    returned=data.frame(
      alpha_ind=summary(tempmodel)$coefficients[1],
      beta_ind=summary(tempmodel)$coefficients[2],
      delta_ind=summary(tempmodel)$coefficients[3]
    )
  }
  return(returned)
  })
##JOIN THE COEFFICIENT DATAFRAME WITH THE DATASET##
finaldata=left_join(finaldata, individnlscoef, by="subject_id")
```
TRANSFORM DISCOUNT FACTOR TO MONTHLY DISCOUNT RATE
```{r non-linear least squares individual parameter estimation three}
finaldata$rate_ind = finaldata$delta_ind^(-30) - 1
```

# Try out nlme

An example:

```{r nlme test, eval=F}
str(Orange)
GGally::ggpairs(Orange)

summary(Orange$circumference)

nlMLMod <- nlme(circumference ~ phi1 / (1 + exp(-(age - phi2)/phi3)),
                data = Orange,
                fixed = phi1 + phi2 + phi3 ~ 1,
                random = pdDiag(phi1 + phi2 ~ 1),
                groups = ~ Tree,
                start = c(phi1 = 200, phi2 = 700, phi3 = 350))

summary(nlMLMod)
coef(nlMLMod)
```


```{r nlme estimate dd}
##Start from the nls model
a_start=0.8027
b_start=0.8402
d_start=0.997

choice_nlme_mod_noa=nlme(sooner_choice ~ ( endowment_later * (b^t0 * d^k * pratio)^(1/(a-1)) ) / ( 1+pratio * (b^t0 * d^k * pratio)^(1/(a-1)) ), 
              data = finaldata, 
              fixed = a + b + d ~ 1,
              random = pdDiag(b + d ~ 1), # we don't model a as a RE because we care less about it
              groups = ~ subject_id,
              method = 'REML',
              start = c(a=a_start,b=b_start, d=d_start),
              control = nlmeControl(pnlsTol=.0225, msVerbose = T))

choice_nlme_mod=nlme(sooner_choice ~ ( endowment_later * (b^t0 * d^k * pratio)^(1/(a-1)) ) / ( 1+pratio * (b^t0 * d^k * pratio)^(1/(a-1)) ), 
              data = finaldata, 
              fixed = a + b + d ~ 1,
              random = pdDiag(a + b + d ~ 1), # we don't model a as a RE because we care less about it
              groups = ~ subject_id,
              method = 'REML',
              start = c(a=a_start,b=b_start, d=d_start),
              control = nlmeControl(pnlsTol=.60475, msVerbose = T))
BIC(choice_nlme_mod_noa)-BIC(choice_nlme_mod)
exp((BIC(choice_nlme_mod_noa)-BIC(choice_nlme_mod))/2)

summary(choice_nlme_mod)

#Or use start from nlsList
choice_nlslist <- nlsList(sooner_choice ~ ( endowment_later * (b^t0 * d^k * pratio)^(1/(a-1)) ) / ( 1+pratio * (b^t0 * d^k * pratio)^(1/(a-1)) ) | subject_id, 
                          data = finaldata[,-c(26:34)], 
                          start = c(a=a_start,b=b_start, d=d_start),
                          control = list(maxiter = 1000, minFactor = 1/2^20))

summary(choice_nlslist, pool = T)

choice_nlme_nlslist <- nlme.nlsList(model = choice_nlslist,
                                   random = pdDiag(a + b + d ~ 1),
                                   method = 'REML',
                                   start = c(a=a_start,b=b_start, d=d_start),
                                   control = nlmeControl(pnlsTol=.33775, msVerbose = T))

BIC(choice_nlme_mod)-BIC(choice_nlme_nlslist)
exp((BIC(choice_nlme_mod)-BIC(choice_nlme_nlslist))/2)

choice_nlme_nlslist_noa <- nlme.nlsList(model = choice_nlslist,
                                   random = pdDiag(b + d ~ 1),
                                   method = 'REML',
                                   start = c(a=a_start,b=b_start, d=d_start),
                                   control = nlmeControl(pnlsTol=.05375, msVerbose = T))

BIC(choice_nlme_nlslist_noa)-BIC(choice_nlme_nlslist)
exp((BIC(choice_nlme_nlslist_noa)-BIC(choice_nlme_nlslist))/2)

choice_nlme_nlslist_ML <- update(choice_nlme_nlslist, method = 'ML')

summary(choice_nlme_nlslist)
round(sapply(coef(choice_nlme_nlslist)-coef(choice_nlme_nlslist_ML),mean),5)
```

After all that, we should choose to go with the `nlme` model fit with the help of nlsList, and that includes a random effect of $\alpha$.

```{r plot individual data, fig.width=40, fig.height=40, cache=T}
#install.packages("viridis")
library(viridis)
library(ggplot2)
finaldata %>%
  mutate(k = factor(k)) %>%
  # filter(subject_id < 50) %>%
ggplot(aes(x = pratio, y = sooner_choice, group = k, color = k))+
  geom_point(alpha=.8, position = position_jitter())+
  geom_line(stat = 'smooth', method = 'gam', formula = y~s(x,k=4,bs='cs'))+
  facet_wrap(~subject_id)+
  scale_color_viridis(discrete = T, begin=0, end=.75)+
  coord_cartesian(y=c(-.5,20.5))+
  theme(panel.background = element_blank())

finaldata %>%
  # filter(subject_id < 50) %>%
ggplot(aes(x = k, y = sooner_choice))+
  geom_point(alpha=.8, position = position_jitter())+
  geom_line(stat = 'smooth', method = 'gam', formula = y~s(x,k=3,bs='cs'))+
  facet_wrap(~subject_id)+
  scale_color_viridis(discrete = T, begin=0, end=.75)+
  coord_cartesian(y=c(-.5,20.5))+
  theme(panel.background = element_blank())

finaldata_pred <- finaldata

finaldata_pred$predicted <- predict(choice_nlme_nlslist)
finaldata_pred %>%
  mutate(k = factor(k)) %>%
  # filter(subject_id < 50) %>%
ggplot(aes(x = pratio, group = k, color = k))+
  geom_point(aes(y = sooner_choice), alpha=.8, position = position_jitter())+
  geom_line(aes(y = predicted))+
  facet_wrap(~subject_id)+
  scale_color_viridis(discrete = T, begin=0, end=.75)+
  coord_cartesian(y=c(-.5,20.5))+
  theme(panel.background = element_blank())
```

```{r extract parameter estimates}
coefs_nlme <- coef(choice_nlme_mod)

coefs_nlme$subject_id <- rownames(coefs_nlme)

coefs_both <- finaldata %>% 
  select(subject_id, alpha_ind, beta_ind, delta_ind) %>%
  distinct(subject_id, alpha_ind, beta_ind, delta_ind) %>%
  mutate(subject_id = as.character(subject_id)) %>%
  left_join(coefs_nlme)
```

```{r}
plot(coefs_both$alpha_ind, coefs_both$a)
plot(coefs_both$beta_ind, coefs_both$b)
ggplot2::qplot(coefs_both$beta_ind, coefs_both$b)+ggplot2::coord_cartesian(y=c(.5,1.4),x=c(.5,1.4))
plot(coefs_both$delta_ind, coefs_both$d)
ggplot2::qplot(coefs_both$delta_ind, coefs_both$d)+ggplot2::coord_cartesian(y=c(.9900,1.005),x=c(.965,1.01005))

```